% !TEX root = ../main.tex

\begin{myProblem}\label{OverarchingProblem}
\textbf{(High-level Behavior Synthesis):}
	Given a system $\mathcal{S}$, which comprises a robot and its primitive capabilities $\mathcal{C}$, a system designer-specified reaction to failures $\mathcal{F}$, and a user-provided task $\mathcal{T}$, in terms of its goals $\mathcal{G}$ and initial conditions\footnote{If the task specification process is taking place online, i.e., during system operation, then the initial conditions could also be detected automatically.}
	 $\mathcal{I}$, automatically generate the software implementation of a high-level plan that \emph{guarantees} the system will either:
	\begin{enumerate}[i]
		\item achieve the goals $\mathcal{G}$ safely, if possible,
		\item or react according to the specification $\mathcal{F}$, if the execution of any primitive capabilities $\mathcal{C}$ does not succeed.
	\end{enumerate}
\end{myProblem}

Problem \ref{OverarchingProblem} involves three main considerations.
First, we have to accept input from non-expert users, in the form of $\mathcal{T}$.
In addition, we have to formally specify the task to be carried out by the system in order to capture the goals and initial conditions, system-specific safety requirements, and the possibility of failure.
This formal specification should also account for the desired system's reaction to failure $\mathcal{F}$.
Finally, we want to automatically generate a high-level plan that is verifiably-correct with respect to the formal specification, as well as its software implementation.
These considerations give rise to the following subproblem statements, which together outline our approach to tackling Problem \ref{OverarchingProblem} for Team ViGIR's Atlas humanoid robot.

\begin{myProblem}\label{DiscreteAbstractionProblem}
\textbf{(Discrete Abstraction):}
Given Atlas' control mode transition constraints and the available actions $\mathcal{A}$, define a discrete abstraction $\mathcal{D}$ of the robot-plus-software system, $\mathcal{S}$, that captures both the execution and possible outcomes of its primitive capabilities (control mode transitions and actions).
In addition, maintain a mapping, $\gamma: \mathcal{D} \rightarrow \mathcal{C}$, which grounds\footnote{We refer the reader to \cite{Grounding2013} for a review of symbol grounding in robotics.}
 the abstract symbols of the discrete abstraction to the primitive capabilities $\mathcal{C}$ of the system $\mathcal{S}$.
\end{myProblem}

\begin{myProblem}\label{SpecificationProblem}
\textbf{(Formal Task Specification):}
Given a task $\mathcal{T}$ in terms of goals $\mathcal{G} = \{ g_1, g_2, \ldots, g_n \}$, the task's initial conditions $\mathcal{I} = \{ i_1, i_2, \ldots, i_m \}$,
the desired reaction to failures $\mathcal{F}$,
and the discrete abstraction $\mathcal{D}$ of the system $\mathcal{S}$ that is to carry out the task,\footnote{In general, a discrete abstraction of the robot's workspace would also be an input to this problem. However, we are not modeling it explicitly here.}
 automatically generate a specification $\mathcal{T}_\mathcal{S}$ that encodes, in a formal language, the task being carried out by $\mathcal{S}$.
\end{myProblem}

To illustrate these concepts, consider again the scenario in Example \ref{Ex:PickupObject}. 
We could say that the task's goals $\mathcal{G}$ are $\{ \mathtt{turn\_valve} \}$ and its initial conditions $\mathcal{I}$ are, e.g., $\{ \mathtt{stand} \}$.
The system $\mathcal{S}$ is Atlas running Team ViGIR's software, as described in Section \ref{S:TeamViGIR}.
Its primitive capabilities $\mathcal{C}$ include walking, rotating the wrist joint, and closing the fingers.
The system-specific safety requirements (e.g., action preconditions and control mode changes) are encoded in $\mathcal{D}$.

\begin{myProblem}\label{BehaviorSynthesisProblem}
\textbf{(Mission Plan \& Software Implementation):}
%\textbf{(Synthesis):}
Given a formal task specification $\mathcal{T}_\mathcal{S}$, synthesize a reactive mission plan that is guaranteed to satisfy $\mathcal{T}_\mathcal{S}$, if one exists.
Additionally, if such a plan exists, and given the mapping $\gamma$, automatically generate a software implementation of the reactive mission plan. 
\end{myProblem}

We present our approach to Problems \ref{DiscreteAbstractionProblem}, \ref{SpecificationProblem}, and \ref{BehaviorSynthesisProblem} in Sections \ref{S:abstraction}, \ref{S:ltl}, and \ref{S:synthesis}, respectively.

% END